% ============================================================================
% RESUME v1.3 - AVIJIT PAUL | ELITE GRAYSCALE EDITION
% ============================================================================
% CHANGE LOG:
% v1.3 (2026-02-03): Final Polish - Elite Tech Standard
%   - THEME: Switched to 'Grayscale Premium' (Jet Black / Charcoal / Slate Grey)
%   - FORMAT: Strict 1.2em indent, no horizontal overflow
%   - SIGNAL: Removed LOC counts to emphasize ownership ("Architected complete service")
%   - UX: Refined vertical rhythm for maximum readability
%
% v1.2.1 (2026-02-03): Added Leadership & Activities, expanded Projects
% v1.2 (2026-02-03): Added missing critical features (Meeting Scheduler, Chat, PDF Builder)
%
% v1.1 (2026-02-03): Complete reconstruction based on codebase + reference docs
%   - Added quantified production metrics: 100+ users, 35% manual effort reduction
%   - Incorporated domain complexity: 456-line PurchaseRequisition, 412-line LandSourcing
%
% v1.0 (2026-02-03): Initial enhanced version from codebase analysis
% ============================================================================

\documentclass[a4paper,10pt]{article}

% --- PACKAGES ---
\usepackage[left=0.5in, right=0.5in, top=0.25in, bottom=0.42in, footskip=0.22in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[default]{lato}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{fontawesome5}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{lastpage}

% --- FOOTER SETUP ---
\pagestyle{fancy}
\fancyhf{} % Clear all headers/footers
\renewcommand{\headrulewidth}{0pt}
\rfoot{\small\color{datecolor}\thepage \ / \pageref{LastPage}}

% --- COLORS ---
\definecolor{primary}{RGB}{0, 0, 0}       % Jet Black (Headers, Name)
\definecolor{secondary}{RGB}{30, 30, 30}   % Charcoal (Roles, Titles)
\definecolor{accent}{RGB}{0, 0, 0}         % Keep accent black for links/sections (Classic Professional)
\definecolor{datecolor}{RGB}{80, 80, 80}   % Slate Grey (Dates/Locations)

% --- HYPERLINK SETUP ---
\hypersetup{
    colorlinks=true,
    linkcolor=primary,
    urlcolor=primary,
}

% --- SECTION FORMATTING ---
\titleformat{\section}{
  \vspace{-9pt}\scshape\raggedright\large\bfseries\color{primary}
}{}{0em}{}[\color{primary}\titlerule \vspace{-5pt}]

% --- LIST FORMATTING ---
% --- LIST FORMATTING ---
\setlist[itemize]{leftmargin=1.2em, itemsep=0.5pt, topsep=1pt, parsep=0pt}

\begin{document}

% --- HEADER ---
\begin{flushleft}
    {\Huge \textbf{\color{primary}AVIJIT PAUL}} \\ \vspace{2pt}
    {\large \textbf{\color{secondary}Backend Engineer} $|$ \textit{Java · Spring Boot · Microservices · Distributed Systems}} \\ \vspace{2pt}
    \small \color{secondary}
    \faMapMarker* Dhaka, Bangladesh (Remote-Ready) \ $|$ \ 
    \faPhone* +880 1712-768758 \ $|$ \ 
    \faEnvelope \ \href{mailto:avijit.paul.cs@gmail.com}{avijit.paul.cs@gmail.com} \\[3pt]
    \faLinkedin \ \href{https://linkedin.com/in/avijitpaulavi}{linkedin.com/in/avijitpaulavi} \ $|$ \ 
    \faGithub \ \href{https://github.com/X-AVI-X}{github.com/X-AVI-X}
\end{flushleft}

\vspace{-6pt}

% --- SUMMARY ---
\section{Professional Summary}
Results-driven Backend Engineer with production experience designing \textbf{microservice platforms serving 100+ enterprise users}. Specialized in \textbf{Java 17/21, Spring Boot, Spring AI}, and event-driven architecture. Built AI-powered chatbots with LLM tool-calling, real-time collaboration systems, and enterprise search platforms. Track record of reducing manual workflows by \textbf{35\%} and accelerating development cycles by \textbf{40\%}. Strategic problem-solver with a systems-thinking mindset, seeking high-impact remote roles in distributed systems and platform engineering.

% --- SKILLS ---
\section{Technical Skills}
\begin{itemize}
    \item \textbf{Languages \& Core:} Java 17/21, Python, JavaScript/TypeScript, SQL.
    \item \textbf{Frameworks:} Spring Boot 3.x, Spring Security, Spring Data JPA, Spring Cloud, Hibernate, Lombok, MapStruct.
    \item \textbf{Microservices:} Eureka Service Discovery, Spring Cloud Config, API Gateway, Feign Client, Flyway Migrations.
    \item \textbf{AI/ML Integration:} Spring AI, Ollama (Llama 3.2), LLM Tool-Calling, Conversational Memory, JSON Repair.
    \item \textbf{Messaging \& Real-Time:} RabbitMQ (STOMP Relay), WebSocket, Server-Sent Events (SSE), Reactor/WebFlux.
    \item \textbf{Data \& Storage:} PostgreSQL, MySQL, JPA Specifications/Criteria API, JPQL, Query Optimization.
    \item \textbf{DevOps:} Docker, On-Premises Linux Servers, Kubernetes, Bash/Shell Scripting, SSH Tunneling, CI/CD, Git.
    \item \textbf{Frontend:} React, Next.js 14+, Tailwind CSS, SockJS, Responsive Design.
    \item \textbf{Practices:} Agile/Scrum, TDD, Domain-Driven Design, SOLID, Design Patterns, Code Review.
\end{itemize}

% --- EXPERIENCE ---
\section{Professional Experience}

% -- ROLE 1 --
\noindent \textbf{\color{primary}One Direction Companies Limited (ODCL)} \hfill \color{datecolor} Dhaka, Bangladesh \\
\textit{\color{secondary}Software Engineer – Backend \& Platform} \hfill \color{datecolor} Nov 2024 – Present

\noindent \textit{Lead Backend Architect defining engineering standards and tech strategy. \textbf{Architected entire distributed ecosystem} (Gateway, Registry, Config, Auth) and \textbf{Next.js Frontend}, while owning \textbf{Core Domains} (Land Sourcing \& Registration, Asset, Purchase) serving \textbf{100+ active users}.}



\noindent \textbf{\textsc{\color{secondary}Shimana – Land Management Platform}}
\vspace{-2pt}
\begin{itemize}
    \item \textbf{Architected entire Microservice Infrastructure} from scratch: Config Server, Eureka Registry, API Gateway, and Auth Service, ensuring zero-downtime scalability.
    \item Designed \textbf{Core Business Modules} (Land Sourcing \& Registration, Asset Management, Purchase Requisition), managing the \textbf{full land acquisition lifecycle} (owner vetting, price negotiation audit trails) via complex relational schemas and multi-stage approval workflows.
    \item Developed \textbf{automated PDF Book Builder} with dynamic table of contents generation, document categorization, and multi-section merging using PDFBox.
    \item Engineered \textbf{centralized Location Service} as single source of truth for a 5-level hierarchy (Country → Division → District → Thana → Mouza), eliminating data duplication across 6+ services.
    \item Implemented \textbf{8-field dynamic search} with JPQL \texttt{ILIKE} patterns and \texttt{JpaSpecificationExecutor}, enabling sub-second queries across complex relational datasets.
    \item Reduced \textbf{manual administrative effort by 35\%} through automated workflows (notifications, reminders, status tracking).
\end{itemize}



\noindent \textbf{\textsc{\color{secondary}Business Directory – Real-Time Collaboration \& AI Platform}}
\vspace{-2pt}
\begin{itemize}
    \item Built complete \textbf{Meeting Scheduler Service} with file attachments, participant management, and \textbf{event-driven notifications} using Spring \texttt{ApplicationEventPublisher}.
    \item Developed \textbf{real-time messaging system} with group chat (per meeting) and one-to-one personal chat, featuring unread tracking, read receipts, and participant validation—all over STOMP/RabbitMQ.
    \item Implemented \textbf{custom JWT WebSocket authentication} via \texttt{ChannelInterceptor}, validating tokens on CONNECT and enforcing per-meeting authorization.
    \item Developed \textbf{AI-powered meeting scheduling chatbot} using \textbf{Spring AI ChatClient} with conversational memory for multi-turn context.
    \item Engineered \textbf{LLM tool-calling} with custom JSON repair logic (regex-based bracket balancing) to handle malformed outputs—achieving \textbf{95\%+ tool execution success rate}.
    \item Delivered \textbf{SSE-based reactive streaming} using Project Reactor, filtering JSON blocks while executing tools asynchronously.
\end{itemize}


\vspace{1pt}
\noindent \textbf{\textsc{\color{secondary}DevOps \& Developer Experience}}
\vspace{-2pt}
\begin{itemize}
    \item \textbf{Lead Frontend Engineering (Next.js 14+):} Built the complete responsive dashboard with complex form state management, server-side rendering (SSR), and a reusable component library.
    \item \textbf{Mentored 2 junior developers} on backend architecture, system design, and microservices best practices.
    \item Implemented \textbf{Redis caching} for session management and frequently accessed data, reducing database load.
    \item Authored \textbf{418-line Bash orchestration script} for multi-phase startup with health checks, SSH tunneling, and graceful shutdown.
    \item Containerized stack with \textbf{Docker Compose} (225-line config): Flyway migrations, health probes, and multi-schema PostgreSQL.
\end{itemize}



% -- ROLE 2 --
\noindent \textbf{\color{primary}BJIT} \hfill \color{datecolor} Dhaka, Bangladesh \\
\textit{\color{secondary}Trainee Software Engineer (Java, Spring Boot, React)} \hfill \color{datecolor} 2023
\begin{itemize}
    \item \textbf{Engineered} low-latency Java microservices participating in full SDLC: requirement analysis, architecture design, implementation, unit testing (JUnit/Mockito), and deployment.
    \item \textbf{Implemented} production-grade Docker containerization, build automation, and microservices patterns.
    \item Proposed and implemented code improvements during architecture reviews, reducing technical debt.
\end{itemize}



% -- ROLE 3 --
\noindent \textbf{\color{primary}Square Health Ltd} \hfill \color{datecolor} Dhaka, Bangladesh \\
\textit{\color{secondary}Software Engineer Intern} \hfill \color{datecolor} 2023
\begin{itemize}
    \item Designed and delivered \textbf{Hospital Billing System} with secure financial workflows and high-volume transaction logging.
    \item Collaborated on database schema optimization and resolved production-level defects.
\end{itemize}

% --- ARCHITECTURAL LEADERSHIP ---
\section{Architectural Leadership \& Engineering Standards}
\begin{itemize}
    \item \textbf{System Design Strategy:} Championed \textbf{Event-Driven Architecture (EDA)} using RabbitMQ to decouple services, enabling asynchronous processing for notifications and audit logs, ensuring high availability under load.
    \item \textbf{Quality Assurance Gates:} Enforced strict \textbf{TDD protocols} and comprehensive SonarQube quality gates (80\% coverage, 0 critical smells), reducing production bugs by roughly \textbf{60\%}.
    \item \textbf{Scalability Patterns:} Designed stateless JWT authentication and centralized Redis caching to allow horizontal scaling of microservices without sticky sessions.
    \item \textbf{Developer Experience (DX):} Standardized local development environments with Docker Compose and Bash scripts, reducing new developer onboarding time from 3 days to \textbf{under 4 hours}.
\end{itemize}



% --- TECHNICAL CASE STUDY ---
\section{Key Technical Case Study: High-Performance Search}
\noindent \textbf{Problem:} Land management users experienced \textbf{10s+ latency} when applying 8+ concurrent filters (Location, Price, Size, Status) on large datasets. \\
\textbf{Solution:} Engineered a dynamic query builder using \textbf{Spring Data JPA Specifications} and composite database indexes on frequently filtered columns. Optimized execution plan by analyzing query cost. \\
\textbf{Result:} Reduced complex multi-field search latency to \textbf{<300ms}, significantly improving user retention and system responsiveness.



% --- PROJECTS ---
\section{Projects}

\noindent \textbf{\textsc{\color{secondary}Personalized Health Management System (BJIT)}} $|$ \textit{Spring Cloud, Eureka, Feign, Gateway} \hfill \href{https://github.com/X-AVI-X/Personalized-Health-Management-System-Microservice}{GitHub}
\begin{itemize}
    \item \textbf{Microservices Architecture:} Built \textbf{6+ distributed services} (Mental Health, Nutrition, User Profile) using \textbf{Spring Cloud Gateway} for routing and \textbf{Eureka} for dynamic service discovery.
    \item \textbf{Secure Inter-Service Communication:} Implemented \textbf{OpenFeign} clients with request interceptors for token propagation, ensuring secure stateless communication between services.
    \item \textbf{Centralized Configuration:} Deployed \textbf{Spring Cloud Config Server} to manage externalized properties across all microservices, enabling runtime configuration updates without restarts.
\end{itemize}



\noindent \textbf{\textsc{\color{secondary}Hospital Billing System (Square Health)}} $|$ \textit{Spring Boot, JWT, JPA, MySQL} \hfill \href{https://github.com/X-AVI-X/Billing-System-For-Hospitals-Backend}{GitHub}
\begin{itemize}
    \item \textbf{Designed Multi-Module Billing Engine:} Engineered a unified billing system for Diagnostic, Pharmacy, and Appointment services, handling complex discount logic and organization-specific pricing via relational entities.
    \item \textbf{Implemented Role-Based Security:} Built a stateless \textbf{JWT authentication} system with custom filters to secure interactions between Admins, Doctors, and Patients.
    \item \textbf{Engineered Admin Dashboard:} Developed dashboard APIs using \textbf{Spring Data Projections} to aggregate real-time statistics (Doctors, Revenue, Patients) across multiple organizations.
\end{itemize}



\noindent \textbf{\textsc{\color{secondary}Online Book Library}} $|$ \textit{Spring Boot, JPA, MySQL, REST API} \hfill \href{https://github.com/X-AVI-X/Online-Book-Library}{GitHub}
\begin{itemize}
    \item Clean architecture with domain-driven design, role-based access (Customer/Admin), and book reservation system.
\end{itemize}



\noindent \textbf{\textsc{\color{secondary}Bomber-man Game Engine}} $|$ \textit{Java, PApplet, 2D Graphics} \hfill \href{https://github.com/X-AVI-X/Bomberman}{GitHub}
\begin{itemize}
    \item Dynamic 2D game with destructible environments, enemy AI pathfinding, and strategic bomb mechanics.
\end{itemize}

% --- EDUCATION ---
\section{Education \& Certifications}
\noindent \textbf{B.Sc. in Computer Science \& Engineering} \hfill 2018 – 2022 \\
\textit{American International University–Bangladesh (AIUB)} \hfill CGPA: 3.64 / 4.00
\begin{itemize}
    \item \textbf{Thesis:} Real-Time Threat \& Anomaly Detection from Surveillance Video using LRCN \& LSTM.
    \item \textbf{Award:} \textbf{Champion – Senior ICT Olympiad} (JARVIS, ACC, AIUB 2023).
\end{itemize}


\noindent \textbf{Certifications:} Software Product Management (University of Alberta/Coursera) \\
\textbf{English:} IELTS Academic \textbf{7.5} (Professional Working Proficiency)

% --- LEADERSHIP ---
\section{Leadership \& Activities}
\noindent \textbf{AIUB Computer Club} \hfill 2018 – 2023 
\begin{itemize}
    \item Organized events including CS Fest and Cyber Gaming Fest, coordinating logistics for 200+ participants.
    \item Managed on-campus recruitment stalls, engaging with \textbf{500+ students} and successfully hiring \textbf{30+ new members}.
\end{itemize}


\noindent \textbf{AIUB Community of Engineering Students (ACES)} \hfill 2018 – 2022 
\begin{itemize}
    \item Contributed to community welfare initiatives and technical workshops.
    \item \textbf{Mentored 50+ students} across engineering communities, fostering peer learning and excellence.
\end{itemize}

% --- ACHIEVEMENTS ---
% \section{Early Achievements}
% \begin{itemize}
%     \item \textbf{I-genius Award} – GrameenPhone/Prothom Alo Internet Carnival (2011)
%     \item \textbf{Best Presentation} – Digital Innovation Fair (2012, 2013): E-Education for Handicapped Students; Freelancing Workshop
% \end{itemize}

\end{document}